---
title: "`{cnd} 0.2.0` release"
subtitle: "A package to support conditions"
date: "2026-10-07"
categories: ["R", "{cnd}"]
draft: true
---

Finally on the second minor relase of `{cnd}`, which brings a few nice small features and a few _fixes_.
This release was motivated by the soon inclusion of `{cnd}` in [`{mark}`](https://jmbarbone.github.io/mark/).

```{r}
library(cnd)
```

## Class name updates

This release introduces a **breaking change** around how class names are generated when a package is included.

Previously, package and class name were separate by a single colon `:`.

    ```r
    class(condition("foo", package = "bar")())[1L]
    #> [1] "bar:foo"
    ```

Now, double colons are used:

    ```r
    class(condition("foo", package = "bar")())[1L]
    #> [1] "bar::foo"
    ```

This also affects all conditions used within `{cnd}`, as well as other class names which contain the `cnd` prefix.

## Defaults

Like the concurrent updates in [`{fuj} v0.3.0`](https://jmbarbone.github.io/posts/fuj-0.3.0/), default classes are now included, without a `package` prefix.

To repeat the example from `{fuj}`:

```{r}
some_letter <- function(x) {
  if (!is.character(x)) {
    stop(type_error("x must be a character"))
  }
  x
}
some_letter("a")
try(some_letter(1))
```

But `{cnd}` contains some extra features.
These default conditions can utilize default error formatting:

```{r}
some_letter <- function(x) {
  if (!is.character(x)) {
    stop(type_error(x = x, expected = character()))
  }
  x
}
try(some_letter(1))
```

## `name = ` over `class = `

A small deprecation introduced will warn against using `condition(class = )`:

```{r}
condition(class = "foo")
```

Originally, this was to accommodate [`{S7}`](https://rconsortium.github.io/S7/)'s soon inclusion and handling of `:=`.

Here's a simplified implimentation of how this might work:

```{r}
`:=` <- function(sym, val) {
  sc <- sys.call()                         # lazily retrieve 'sym' and 'val' from call
  sc[[1L]] <- quote(`<-`)                  # replace `:=` with `<-`
  sc[[3L]]$name <- as.character(sc[[2L]])  # redirected to 'name' param
  Exec(sc, parent.frame())                 # Exec() evaluates without including an additional stack
}

Foo := condition() # identical to `Foo <- condition(name = "Foo")`
Foo
```

Perhaps a future update might include the _walrus_ operator in `{cnd}` and/or `{fuj}`, or it might be available through base **R**.
